---
title: "Lab 06 - Ugly charts and Simpson's paradox"
author: "Sophie Boyd"
date: "2-26-26"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox)
library(mosaicData) 
```

### Instructional Staff Employment Trends

```{r read-staff-data}
staff <- read_csv("data/instructional-staff.csv")
```

```{r wide-to-long}
staff_long <- staff %>%
  pivot_longer(cols = -faculty_type, names_to = "year") %>%
  mutate(value = as.numeric(value))
```

```{r line-graph}
staff_long %>%
  ggplot(aes(x = year, y = value, color = faculty_type)) +
  geom_line()
```

```{r line-graph-group}
staff_long %>%
  ggplot(aes(
    x = year,
    y = value,
    group = faculty_type,
    color = faculty_type
  )) +
  labs (x = "Year",
        y = "Percentage of Hires",
        title = "Percentage of Hires by Position (1975 to 2011)") +
  geom_line()
```

If the objective of this plot was to show that the proportion of part-time faculty had gone up compared to other positions, I would make all of the lines one color except for the part-time faculty line to make it stand out. I would also change the title to highlight the comparison between part-time hires and other positions.

### Exercise 2

Remove this text, and add your answer for Exercise 1 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 3

...

Add exercise headings as needed.
